---
title: 'Lab 1: Wrangling Review'
author: "Matthieu Huy"
date: "2023-01-13"
output: html_document
---

```{r setup, include = TRUE, echo = TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(sf)
library(tmap)
```

## Read in data

```{r}
sf_trees <- read_csv(here("data/sf_trees/sf_trees.csv"))
```

##Part 1: Wrangling and ggplot review
####Example 1: Find counts of observations by legal_status and wrangle

```{r}
top_5_status <- sf_trees |>
  count(legal_status) |> #get a count of the all values in legal status column
  drop_na(legal_status) |> #drop any na values
  rename(tree_count = n) |> #rename "n" to "tree_count"
  relocate(tree_count) |> #move tree_count column to front
  slice_max(tree_count, n = 5) |> #keep only the 5 most common legal statuses
  arrange(-tree_count) #rearrange in descending order

ggplot(data = top_5_status,
       aes(x = fct_reorder(legal_status, tree_count), #reorder legal statuses by count
           y = tree_count)) + 
  geom_col() +
  labs(y = "Tree Count",
       x = "Legal Status") +
  coord_flip() +
  theme_minimal()
```

#### Example 2: Only keep observations where legal status is Permitted Site and caretaker is MTA. Store as permitted_mta.

```{r}
#only select rows where legal status and caretaker columns match "Permitted Site" and "MTA"
permitted_mta <- sf_trees |>
  filter(legal_status == "Permitted Site", caretaker =="MTA") 
```

#### Example 3: Only keep Blackwood Acacia trees, then only keep columns legal_status, date, latitude and longitude. Store as blackwood_acacia.

```{r}
blackwood_acacia <- sf_trees |>
  filter(str_detect(species, "Blackwood Acacia")) |>
  select(legal_status, date, latitude, longitude)

# Make a little graph of locations (note R doesn't know these are spatial)

ggplot(data = blackwood_acacia,
       aes(x = longitude,
           y = latitude)) +
  geom_point()
```

#### Example 4: Meet tidyr::separate(). Separate the species column into two separate columns: spp_scientific and spp_common

```{r}
sf_trees_sep <- sf_trees |>
  separate(species, into = c("spp_scientific", "spp_comon"), sep = "::")
```

#### Example 5: Meet tidyr::unite(). Yeah, it does the opposite. Unite the tree_id and legal_status columns, using a separator of “COOL” (no, you’d never actually do this…).

```{r}
sf_trees_unite <- sf_trees |>
  unite("id_status", tree_id:legal_status, sep = "_COOL_")
```

##Part 2: Make some actual maps

